---
title: "Forecast daily bike rental demand using time series models"
date: "r Sys.Date"
output: github_document
author: Natalia Miller
---

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on forecasting daily bike rental demand using time series models in R. It contains analysis such as data exploration, summary statistics and building the time series models. The final report was completed on `r date()`. 

**Data Description:**

This dataset contains the daily count of rental bike transactions between years 2011 and 2012 in Capital bikeshare system with the corresponding weather and seasonal information.

**Data Source:** https://archive.ics.uci.edu/ml/datasets/bike+sharing+dataset

**Relevant Paper:** 

Fanaee-T, Hadi, and Gama, Joao. Event labeling combining ensemble detectors and background knowledge, Progress in Artificial Intelligence (2013): pp. 1-15, Springer Berlin Heidelberg



# Task One: Load and explore the data

## Load data and install packages

```{r}
install.packages("timetk")
install.packages("tidyverse")
install.packages("skimr")
install.packages("dplyr")
install.packages("plyr")
install.packages("lubridate")
library(timetk)
library(tidyverse)
library(skimr)
library(dplyr)
library(plyr)
library(lubridate)
data("bike_sharing_daily")
```


## Describe and explore the data

```{r}

```

Is there a correlation between the normalized temperature and normalized feeling temperature and the total count of bike rentals?
```{r}
#calculate correlation coefficient
cor(bike_sharing_daily$temp, bike_sharing_daily$cnt)
cor(bike_sharing_daily$atemp, bike_sharing_daily$cnt)
```

What are the mean and median temperatures for different seasons (Winter, Fall, Summer, and Spring)?
```{r}
# Calculate mean and median of temperatures for each season
average_temperatures <- bike_sharing_daily %>%
	summarize( winter = mean(temp [season == "1"], na.rm = T),
		fall = mean(temp[season == "4"], na.rm = T),
		summer = mean(temp[season == "3"], na.rm = T),
		spring = mean(temp[season == "2"], na.rm = T))

median_temperatures <- bike_sharing_daily %>%
	summarize( winter = median(temp[season == "1"], na.rm = T),
		fall = median(temp[season == "4"], na.rm = T),
		summer = median(temp[season == "3"], na.rm = T),
		spring = median(temp[season == "2"], na.rm = T))
```

What are the mean temperature, humidity, wind speed, and total rentals per month? 
```{r}
# Create a data frame with mean temperature, humidity, wind speed, 
# and total rentals per month 
monthly_averages <- bike_sharing_daily %>% 
	group_by(month = mnth) %>%
	summarize(avg_temp = mean(temp), avg_humidity = mean(hum),
		avg_wind_speed = mean(windspeed), total_rentals = sum(cnt))
```

Is temperature associated with bike rentals (registered vs. casual)?
```{r}
cor(bike_sharing_daily$temp, bike_sharing_daily$casual)
cor(bike_sharing_daily$atemp, bike_sharing_daily$registered)
```

How do the temperatures change across the seasons?
```{r}
boxplot(average_temperatures, 
	ylim = c(0.3, 0.75), 
	main = "Average Temperature per Season")
```

# Task Two: Create interactive time series plots

```{r}
bike_sharing_daily %>% 
  plot_time_series(dteday, cnt, 
                   .interactive = TRUE,
                   .plotly_slider = TRUE)
```


# Task Three: Smooth time series data

```{r}

```



# Task Four: Decompose and assess the stationarity of time series data

```{r}

```



# Task Five: Fit and forecast time series data using ARIMA models

```{r}

```



# Task Six: Findings and Conclusions
● What did you learn throughout the process?
● Are the results what you expected?
● What are the key findings and takeaways?




























